The NIME of the Title Is Hello

### 2025-01-15 src/article.tex
cancel cmds

The project look at restoring tactile and haptic response of 17th
century harpsichords for an exhibition at the musical instrument museum
San Colombano in Bologna. San Colombano hosts the Tagliavini Collection
(from Luigi Ferdinando Tagliavini), a collection of approximately 70
historical keyboard instruments in working order \cite{Tagliavini2009}
(https://genusbononiae.it/en/san-colombano-tagliavini-collection/). Some
items in the collection are not in playing condition or are too fragile
to continue to be available for the public to play. One of those fragile
instruments in the Tagliavini collection is a 1547 Harpsichord by
Alessandro Trasuntino \cite{Treccani2019}, which was used as the basis
for an interactive exhibition commisioned by San Colombano.

@@

The paper by McAlpine \cite{McAlpine2014} also has very similar
motivations to the project presented here. Both were started from a
musical instrument museum, housing a collection of historic instruments
gifted by a collector (Benton Fletcher / Taglivaini) who wished for
collection to remain playable, which is restricted growing fragility of
some items in the respective collections. Bothe projects center around
the creation of a custioom MIDI controller for interfacing with sound
synthesis engine. The keyboard consisted of two modified commercial MIDI
keyboard controllers in a two manual harpsichord arrangement. Weherease
the McAlpine keyboard used weighted keys, this interface extends the
idea by using the mechanism and strings under tension to provide the
haptic feedback. McAlpine posits a fully paremterised haptic keyboard,
but before taking such a leap we look at what affect a more literal
haptic keyboard may provide to the performer experience.

The previous NIME project ``Tromba Moderna'' \cite{Baldwin2016}, which
looked at reconstructing a full tromba marina to be used as a digital
interface. The one of the limitations flagged by McAlpine was the
oneffectiveness of weighted keys in mimicing the harpischord. The Tromba
Moderna ``attempted to avoid this problem by constructing a fully
functioning replica of the tromba marina'' \cite{Baldwin}. This project
applies the same approach to the harpsichord.m where the projects
deviate is a questions of scale.

@@

Extension to \cite{McAlpine2014} and \cite{Baldwin2016} are the open
sourcing of hardware and software for the project.

@@

Discussion of spatial constraints and affects on electyronics design and
installation is covered in

\section{setup}

@@

The keyboard is fully enclosed and internally has two chambers, front
and rear, in which the electronics could be installed \figure(underside
49-key).

@@

Accomodations were made in the design to allow for as much flexibility
as possible. This also meant that designs for the eletronics would have
to include a deal of flexibility in how they might be installed. Surplus
space was added into the design of the internal chamber of the model
\FIGURE, but there was still a limit given the structural and aesthatic
requirements.

@@

\begin{itemize}
\tightlist
\item
  \cite{HapticKey}
\item
  \cite{mcPherson1} \cite{mcPherson2}
\item
  \cite{Mudd2013}
\item
  \cite{FritzClaudia2017Leon}
\end{itemize}

@@

Using a similalr apporach to @HapticKey the first step was to fabricate
or procure a model of a single working key. San Colombano provided a
model 3-key of a harposchord mechanism by Graziano Bandini \FIGURE. The
3-key model was used a test-bed for potential sensors while designs were
drawn for a full-scale model. The cross-sectional nature of the model
meant it was easy to fit and remove sensors. Small size meant that it
would also be easy to transfer between workshops.

@@

Following from the work in \cite{McPherson1} \cite{McPherson2} we tested
a system using the Fairchild QRE1113 \datasheetfigue. The QRE1113 is a
combination infrared LED and phototransistor sensitive to IR light. The
phototransistor of QRE1113 provides data on how much light is present
and in particular how much light is being relfected from a surface.
Since refelected light will be proportional to the distance of a surfcae
it is a good, close proximity, distance sensor. Distance is the way in
which the instruments in \citeMcpherson1\&2 function as well the Moog
Piano Bar from which it too inspiration.

@@

A rough performance test was later carried out using an STM32
NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
more performative. To execute a single sensor read cycle of the firmware
the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
on average. The Nano ESP32 had high variability so it was decided to
continue with the Nano BLE. Further discussion of using ESP32 continues
in

\section{Going forward}

@@

TLC5940 or other PWM driver such as in \cite{mcpherson} added complexity
to assembling, but were avoid in favour of Addressable LEDs with
intergrated driver. The intergrated LEDs share the same data line which
has the limitation where a fualty LED will stop all following LEDs from
working. An individual PWM driver per PCB would allow avoid situations
where failutre cascades, but would add complexity in assembly as the
parts are typically SMD. Also, given the LEDs are fixed to the board and
could not be repaired in situ, the solution would be to replace the PCB
regardless of which approach was taken. Given the LEDs were only used in
debugging and calibration and otehrwise would not be visible, a simpler
design approach was taken.

@@

Transistor / Op-amp circuit could have been used similar to the one in
\cite{mcpherson}.

@@

A locator was adapted from a creative commons stl model
https://www.thingiverse.com/thing:1646016 that allowed a generic
crimping tool to be adapted for use with JST-PH crimps† \figure{}. The
locator was printed using a Bambu X1 3D printer.

@@

A misreading in the original clearance between the keys and the PCB
meant that the keys wer block entirely from moving. A 10mm by 15mm
section was milled out of each lke \FIGURE to create extra clearance.
The milling required the removal, shortening and refitting of the
leather pads for the jacks \FIGURE. The extra work added an extra day to
the full install.

>>>>

The project look at restoring tactile and haptic response of 17th
century harpsichords for an exhibition at the musical instrument museum
San Colombano in Bologna. San Colombano hosts the Tagliavini Collection
(from Luigi Ferdinando Tagliavini), a collection of approximately 70
historical keyboard instruments in working order
\textbackslash cite\{Tagliavini2009\}
(https://genusbononiae.it/en/san-colombano-tagliavini-collection/). Some
items in the collection are not in playing condition or are too fragile
to continue to be available for the public to play. One of those fragile
instruments in the Tagliavini collection is a 1547 Harpsichord by
Alessandro Trasuntino \textbackslash cite\{Treccani2019\}, which was
used as the basis for an interactive exhibition commisioned by San
Colombano.

@@

The paper by McAlpine \textbackslash cite\{McAlpine2014\} also has very
similar motivations to the project presented here. Both were started
from a musical instrument museum, housing a collection of historic
instruments gifted by a collector (Benton Fletcher / Taglivaini) who
wished for collection to remain playable, which is restricted growing
fragility of some items in the respective collections. Bothe projects
center around the creation of a custioom MIDI controller for interfacing
with sound synthesis engine. The keyboard consisted of two modified
commercial MIDI keyboard controllers in a two manual harpsichord
arrangement. Weherease the McAlpine keyboard used weighted keys, this
interface extends the idea by using the mechanism and strings under
tension to provide the haptic feedback. McAlpine posits a fully
paremterised haptic keyboard, but before taking such a leap we look at
what affect a more literal haptic keyboard may provide to the performer
experience.

The previous NIME project ``Tromba Moderna''
\textbackslash cite\{Baldwin2016\}, which looked at reconstructing a
full tromba marina to be used as a digital interface. The one of the
limitations flagged by McAlpine was the oneffectiveness of weighted keys
in mimicing the harpischord. The Tromba Moderna ``attempted to avoid
this problem by constructing a fully functioning replica of the tromba
marina'' \textbackslash cite\{Baldwin\}. This project applies the same
approach to the harpsichord.m where the projects deviate is a questions
of scale.

@@

Extension to \textbackslash cite\{McAlpine2014\} and
\textbackslash cite\{Baldwin2016\} are the open sourcing of hardware and
software for the project.

@@

Discussion of spatial constraints and affects on electyronics design and
installation is covered in \textbackslash section\{setup\}

@@

The keyboard is fully enclosed and internally has two chambers, front
and rear, in which the electronics could be installed
\textbackslash figure(underside 49-key).

@@

Accomodations were made in the design to allow for as much flexibility
as possible. This also meant that designs for the eletronics would have
to include a deal of flexibility in how they might be installed. Surplus
space was added into the design of the internal chamber of the model
\textbackslash FIGURE, but there was still a limit given the structural
and aesthatic requirements.

@@

\begin{itemize}
\tightlist
\item
  \textbackslash cite\{HapticKey\}
\item
  \textbackslash cite\{mcPherson1\} \textbackslash cite\{mcPherson2\}
\item
  \textbackslash cite\{Mudd2013\}
\item
  \textbackslash cite\{FritzClaudia2017Leon\}
\end{itemize}

@@

Using a similalr apporach to @HapticKey the first step was to fabricate
or procure a model of a single working key. San Colombano provided a
model 3-key of a harposchord mechanism by Graziano Bandini
\textbackslash FIGURE. The 3-key model was used a test-bed for potential
sensors while designs were drawn for a full-scale model. The
cross-sectional nature of the model meant it was easy to fit and remove
sensors. Small size meant that it would also be easy to transfer between
workshops.

@@

Following from the work in \textbackslash cite\{McPherson1\}
\textbackslash cite\{McPherson2\} we tested a system using the Fairchild
QRE1113 \textbackslash datasheetfigue. The QRE1113 is a combination
infrared LED and phototransistor sensitive to IR light. The
phototransistor of QRE1113 provides data on how much light is present
and in particular how much light is being relfected from a surface.
Since refelected light will be proportional to the distance of a surfcae
it is a good, close proximity, distance sensor. Distance is the way in
which the instruments in \textbackslash citeMcpherson1\&2 function as
well the Moog Piano Bar from which it too inspiration.

@@

A rough performance test was later carried out using an STM32
NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
more performative. To execute a single sensor read cycle of the firmware
the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
on average. The Nano ESP32 had high variability so it was decided to
continue with the Nano BLE. Further discussion of using ESP32 continues
in \textbackslash section\{Going forward\}

@@

TLC5940 or other PWM driver such as in \textbackslash cite\{mcpherson\}
added complexity to assembling, but were avoid in favour of Addressable
LEDs with intergrated driver. The intergrated LEDs share the same data
line which has the limitation where a fualty LED will stop all following
LEDs from working. An individual PWM driver per PCB would allow avoid
situations where failutre cascades, but would add complexity in assembly
as the parts are typically SMD. Also, given the LEDs are fixed to the
board and could not be repaired in situ, the solution would be to
replace the PCB regardless of which approach was taken. Given the LEDs
were only used in debugging and calibration and otehrwise would not be
visible, a simpler design approach was taken.

@@

Transistor / Op-amp circuit could have been used similar to the one in
\textbackslash cite\{mcpherson\}.

@@

A locator was adapted from a creative commons stl model
https://www.thingiverse.com/thing:1646016 that allowed a generic
crimping tool to be adapted for use with JST-PH crimps†
\textbackslash figure\{\}. The locator was printed using a Bambu X1 3D
printer.

@@

A misreading in the original clearance between the keys and the PCB
meant that the keys wer block entirely from moving. A 10mm by 15mm
section was milled out of each lke \textbackslash FIGURE to create extra
clearance. The milling required the removal, shortening and refitting of
the leather pads for the jacks \textbackslash FIGURE. The extra work
added an extra day to the full install.
